### 2016-05-24 ms/ms_probaweb.tex
Ready for submission

\textbf{Title:} Bringing Elton and Grinnell together: a quantitative framework to represent the biogeography of ecological interaction networks

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo
Martinez$^{6}$, Tommi Nyman$^{7}$, Timoth\'ee Poisot$^{2,8}$, Spencer A. Wood$^{9}$, Daniel B. Stouffer$^{10}$, Jason M. Tylianakis$^{10,11}$ Tomas Roslin$^{12}$,\\

1: Canada Research Chair in Integrative Ecology. D\'epartement de
biologie, Universit\'e de Sherbrooke, 2500 Boulevard l'Universit\'e,
Sherbrooke (Québec). J1K 2R1\\

2: Qu\'ebec Centre for Biodiversity Sciences\\

3: University of Florida, Department of Wildlife Ecology and Conservation, 110 Newins-Ziegler Hall, PO Box 110430, Gainesville, Fl. 32611-0430 \\

4: Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA.\\

5: Senckenberg Research Institute and Natural History Museum, Senckenberganlage 25, D-60325 Frankfurt am Main, Germany\\

6: Ecology and Evolutionary Biology Department, University of Arizona, P.O. Box 210088, Tucson, AZ 85721, USA\\

7: University of Eastern Finland, Department of Environmental and Biological Sciences, P.O. Box 111, FI-80101 Joensuu, Finland\\

8: Université de Montréal, Département des Sciences Biologiques, 90 Avenue Vincent d’Indy, Montréal, QC H2V3S9, Canada.\\

9: University of Washington, School of Environmental and Forest Sciences,
Box 352100, Seattle, WA 98195, USA\\

10: Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch, New Zealand\\

11: Department of Life Sciences, Imperial College London, Silwood Park Campus, 
Buckhurst Road, Ascot, Berkshire SL5 7PY, United Kingdom\\

12: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, 750 07 Uppsala, Sweden\\

@@

\textbf{Words in the abstract:}

\textbf{Words in the main text:}

\textbf{Words in the legends:}

\textbf{Figures: 6}

\textbf{Tables: 2}

\textbf{References:}

@@

Biogeography has traditionally focused on the spatial distribution and
abundance of species. Both are driven by the way species interact with one
another, but also by the way these interactions vary across time and space.
Here, we call for an integrated approach in which community structure
represented as a network of ecological interactions, and show how it
translates to biogeography questions. We propose that the ecological niche may
be redefined to encompass the effect of the environment on species
distribution (the Grinnellian dimension of the niche) and on the ecological
interactions among them (the Eltonian dimension). Starting from this 
concept, we develop a quantitative theory to explain turnover of interactions
in space and time -– \emph{i.e.} a novel approach to interaction distribution
modelling. We apply this framework to host–parasite interactions across Europe
and find that two aspects of the environment (temperature and precipitation)
exert a strong imprint on species co-occurrence, but not on species
interactions. Even where species co-occur, interaction proves to be stochastic
rather than deterministic, adding to variation in realized network structure.
We also find that a large majority of host-parasite pairs are never found
together, thus precluding any inferences regarding their probability to
interact. This first attempt to explain variation of network
structure at large spatial scales opens new perspectives at the interface
of species distribution modelling and community ecology.

@@

Community ecology is \textit{the study of the interactions that determine
the distribution and abundance of organisms} \citep{Krebs2001}. Despite a
general consensus on this definition (Scheiner2007), research on variation in
community structure has mostly focused on the spatial and temporal turnover
of species composition \citep{Anderson2011}, neglecting variation in the way
species interact with each other, despite accumulating empirical evidence that
this is a major source of diversity \citep{Poisot2015a}. Given this omission,
it is perhaps not surprising that biogeographers are still struggling to
establish whether interactions actually impact the distribution of species at
large spatial scales \citep{Wisz2012, Kissling2012}: treating interactions
as fixed events neglects a large part of the complexity of empirical
communities, and will most likely deliver underwhelming results. Recent
attempts at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013} have brought some methodological advances,
but are not sufficient for two reasons. First, these techniques are still
based on a `species-based' approach to communities, where interactions
are merely treated as fixed covariates affecting distribution. Second,
they failed to provide a conceptual step forward, both in their treatment
of interactions and in the quality of the predictions they make.

Because it is an explicit description of species interactions, network
approaches offer a convenient representation of communities. Species are
represented as nodes, so that networks already encompass all the information
used by current approaches; in addition, interactions are represented by
links, so that networks provide additional, higher-order information on
community structure. To date, studies of network diversity have mostly been
concerned with the distribution of interactions within locations, and less so
with variation among locations \citep{Dunne2005, Bascompte2007, Ings2007,
Kefi2012}. There is, however, ample evidence that interaction networks vary in
space and time \citep{Laliberte2010, Poisot2012, Albouy2014, Poisot2016,
Trojelsgaard2015}, even though there is no common framework with which to
generalize these results. Metacommunity theory provides explanations for
variation in the distribution of the nodes \citep{Gravel2011c, Pillai2011},
but there is no such explanation to the variation of node and link
occurrences. Consequently, we urgently need a conceptual framework to
formalize these observations, as it is the only way towards fulfilling the
goal of community ecology: providing cogent predictions about, and
understanding of, the structure of ecological communities.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a clear need to bring the two together.
Here, we offer an integrated approach to do so, adopting the view that
community structure is best represented as a network of ecological
interactions. Based on this idea, we propose a new description of the basic
concept of the ecological niche that integrates the effect of the environment
on species distribution and on the ecological interactions among them.
Starting from this redefined concept, we develop a quantitative theory to
explain turnover of interactions in space and time. We first present the
conceptual framework, and then formalize it mathematically, using a
probabilistic model to represent the sampling of the regional pool of
interactions. At the level of species pairs, the statistical approach could be
conceived as an interaction distribution model. At the community level, the
approach provides a likelihood-based method to compare different hypotheses of
network turnover. As an illustrative example, we apply this novel framework to
a large data set on host–parasite interactions across Europe and perform test
hypotheses underlying variation of species and interactions. The network
structure changes systematically across the latitudinal gradient, with a peak
of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, or \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized locally, revealing the stochastic nature of
ecological interactions. Our objective here is to provide a conceptual
framework to explain the sampling of the regional pool of interactions, along
with a quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-
occur locally and second why these species do or do not interact.

There are multiple causes of spatial turnover of species co-occurrence. The first
and most-studied driver is the effect of variation in the abiotic environment
on species performance. Combined with specific responses in demography, it
generates variation among sites by selecting the locally fittest species
\citep{Leibold2004}. Stochasticity plays an additional role, either because
colonization and extinction events \citep{Hanski1999} are inherently
unpredictable or because strong non-linear feedbacks in community dynamics
generate alternative transients and equilibria \citep{Chase2007, Vellend2014}.
Analyses of community turnover are usually performed with data represented in
a table with rows corresponding to sites (or measurements) and columns to
species. Metrics of beta diversity quantify the variance of this community
data \citep{Legendre2005}. Traditional approaches rely on measures of
dissimilarity among communities, such as the Jaccard or Bray–Curtis indices.
More recent approaches decompose total variation of the community data into
species and site contributions to beta diversity \citep{Legendre2013}, and
further partition it into dissimilarity due to changes in species richness and
dissimilarity due to actual species turnover \citep{Baselga2010;
Carvalho2012}. Even though these methods compare whole lists of species among
sites or measurements, they remain fundamentally `species-based', since they
report variation within columns. None of them explicitly considers variation
of associations (i.e., of pairs or higher-order motifs –
\citealt{Stouffer2007}).

Similarly, we are getting a better understanding of interaction turnover. As
mentioned above, in the network approach to community structure, species and
interactions are represented by nodes and links, respectively. Associations
can also be represented by matrices in which entries represent the occurrence
or intensity of interactions among species (rows and columns). Network
complexity is then computed as the number of interactions (in the case of
binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition can be independent of variation of ecological
interactions, suggesting that species and interaction distribution may well
respond to different drivers \citep{Poisot2012}.

The "niche" is by far the dominant concept invoked to explain species
distributions and community assembly, from the local to the global scale.
Following \citealt{Hutchinson1957}, the niche is viewed as the set of
environmental conditions allowing a population to establish and persist (see
also \citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to put into practice with species distribution models, as it
maps naturally on available distributional and environmental data. Consequently, a vast array of statistical tools have been developed to
implement it (e.g. BIOMOD \citealt{Thuiller2003}, MaxEnt
\citealt{Phillips2006}). It is however much harder to account for ecological
interactions within this approach \citep{Peterson2011}. As such, these
interactions are often viewed as externalities constraining or expanding the
range of environmental conditions required for a species to maintain a viable
population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

Interestingly, the ecological network literature also has its own "niche
model" to position a species in a community \citep{Williams2000}. The niche of
a species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000, Eklof2013}.
The niche model of food web structure and its variants have successfully
explained the complexity of a variety of networks, from food webs to
plant–pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

Despite several attempts to update the concept of the ecological niche,
ecologists have not moved far beyond the "n-dimensional hypervolume" defined
by Hutchinson. Despite its intuitive interpretation and easy translation into
species distribution models \citep{Boulangeat2012, Blonder2014}, the concept
has been frequently criticized \citep{Hardin1960, Peters1991, Silvertown2004},
and several attempts have been made to expand and improve it
\citep{Pulliam2000, Chase2003, Soberon2007, Holt2009, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}.
The Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. While this aspect of the niche is well
known by community ecologists, it is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000} and its parameterization using functional traits
\citep{Gravel2013; Bartomeus2016} made it more operational.

These perspectives are rather orthogonal to each other, and this has resulted
in considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
"[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions".
Their representation merges zero-net-growth isoclines delimiting the
Grinnellian niche ("when does the population persists?"") with impact vectors
delimiting the Eltonian niche ("what is the per-capita impact?"). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}), it remains
impractical at larger spatial scales because of the difficulties in measuring
it. The absence of any mathematical representation of the niche that can be
easily fit to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions. Thus, a more integrative description of the niche
will be key to understand spatial and temporal turnover in community
structure.

@@

This approach radically changes the representation of the niche, putting species
distributions and ecological interactions into the same formalism.
Moreover, it allows the limits of the niche axes to be independent of each other (as in the
example in Fig. 2), or to interact. For instance, the optimal
prey size for predatory fishes could decline with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. Alternatively, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses concerning temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation, there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
constrain distribution \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent, and only the marginal occurrence must be accounted for, $P(
X_{ijy}}|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976,
Connor1980, Gotelli2000}). Second- and higher-order interactions (e.g. trophic
cascades) could also affect co-occurrence. Co-occurrence of multiple species
embedded in ecological networks is a topic of its own, however, and is
influenced by both network topology and species richness \citep{Cazelles2016}.
Not only direct interactions influence co-occurrence, but indirect
interactions do as well (e.g. plant species sharing an herbivore, or
herbivores sharing parasitoids, could repel each other in space
\citealt{Holt1993}). The impact of direct interactions and first-order
indirect interactions on co-occurrence tends to vanish with increasing species
richness in the community. Further, co-occurrence is also influenced by the
covariance of interacting species to an environmental gradient
\citep{Cazelles2015}. Because of the complexity of relating co-occurrence to
the structure of interaction networks, we will focus here on the variation of
interactions and not on their distribution, and leave this specific issue for
the Perspectives section and future research.

@@

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy} |E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to  be sensitive to the environment. For instance,
\citep{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions also propagate up the
community and influence network structure \citep{Tylianakis2007, Woodward2010;
Petchey2010; Lelong2015}.

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}, and is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.

@@

Understanding the drivers of spatial variation in network structure is a
key problem to solve in order to anticipate the impacts of global changes on
biodiversity. Our representation of spatial variation of community
structure presents a new approach for the study of the biogeography of ecological
networks. We see the following key challenges and opportunities ahead in this
exciting area of research:

>>>>

\textbf{Title:} Bringing Elton and Grinnell together: a quantitative framework
to represent the biogeography of ecological interaction networks\\

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer
A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo D. Martinez$^{6}$, Tommi
Nyman$^{7}$, Timothée Poisot$^{2,8}$, Spencer A. Wood$^{9}$, Daniel B.
Stouffer$^{10}$, Jason M. Tylianakis$^{10,11}$ Tomas Roslin$^{12}$ \\

\textbf{1}: Canada Research Chair in Integrative Ecology. Département de
biologie, Université de Sherbrooke, 2500 Boulevard l'Université,
Sherbrooke (Québec). J1K 2R1\\
\textbf{2}: Québec Centre for Biodiversity Sciences\\
\textbf{3}: University of Florida, Department of Wildlife Ecology and Conservation, 110 Newins-Ziegler Hall, PO Box 110430, Gainesville, Fl. 32611-0430 \\
\textbf{4}: Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA.\\
\textbf{5}: Senckenberg Research Institute and Natural History Museum, Senckenberganlage 25, D-60325 Frankfurt am Main, Germany\\
\textbf{6}: Ecology and Evolutionary Biology Department, University of Arizona, P.O. Box 210088, Tucson, AZ 85721, USA\\
\textbf{7}: University of Eastern Finland, Department of Environmental and Biological Sciences, P.O. Box 111, FI-80101 Joensuu, Finland\\
\textbf{8}: Université de Montréal, Département des Sciences Biologiques, 90 Avenue Vincent d’Indy, Montréal, QC H2V3S9, Canada.\\
\textbf{9}: University of Washington, School of Environmental and Forest Sciences,
Box 352100, Seattle, WA 98195, USA\\
\textbf{10}: Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch, New Zealand\\
\textbf{11}: Department of Life Sciences, Imperial College London, Silwood Park Campus, 
Buckhurst Road, Ascot, Berkshire SL5 7PY, United Kingdom\\
\textbf{12}: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, 750 07 Uppsala, Sweden\\

@@

\textbf{Words in the abstract: 226}

\textbf{Words in the main text: 7043}

\textbf{Words in the legends: 502}

@@

Biogeography has traditionally focused on the spatial distribution and
abundance of species. Both are driven by the way species interact with one
another, but also by the way these interactions vary across time and space.
Here, we call for an integrated approach, adopting the view that community
structure is best represented as a network of ecological interactions, and
show how it translates to biogeography questions. We propose that the
ecological niche should encompass the effect of the environment
on species distribution (the Grinnellian dimension of the niche) and on the
ecological interactions among them (the Eltonian dimension). Starting from
this concept, we develop a quantitative theory to explain turnover of
interactions in space and time -- \emph{i.e.} a novel approach to interaction
distribution modelling. We apply this framework to host–parasite interactions
across Europe and find that two aspects of the environment (temperature and
precipitation) exert a strong imprint on species co-occurrence, but not on
species interactions. Even where species co-occur, interaction proves to be
stochastic rather than deterministic, adding to variation in realized network
structure. We also find that a large majority of host-parasite pairs are never
found together, thus precluding any inferences regarding their probability to
interact. This first attempt to explain variation of network structure at
large spatial scales opens new perspectives at the interface of species
distribution modelling and community ecology.

@@

Community ecology is \textit{the study of the interactions that determine
the distribution and abundance of organisms} \citep{Krebs2001}. Despite a
general consensus on this definition \citep{Scheiner2007a}, research on variation in
community structure has mostly focused on the spatial and temporal turnover
of species composition \citep{Anderson2011}, neglecting variation in the way
species interact with each other despite accumulating empirical evidence that
this is a major source of diversity \citep{Poisot2015a}. Given this omission,
it is perhaps not surprising that biogeographers are still struggling to
establish whether interactions actually impact the distribution of species at
large spatial scales \citep{Wisz2012, Kissling2012}: treating interactions
as fixed events neglects a large part of the complexity of empirical
communities, and will most likely deliver underwhelming results. Recent
attempts at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013} have brought some methodological advances,
but are not sufficient for two reasons. First, these techniques are still
based on a `species-based' approach to communities, where interactions
are merely treated as fixed covariates affecting distribution. Second,
they failed to provide a conceptual step forward, both in their treatment
of interactions and in the quality of the predictions they make.

Network approaches offer a convenient representation of communities because
they simultaneously account for species composition and their interactions.
Species are represented as nodes, so that networks already encompass all the
information used by current approaches; in addition, interactions are
represented by links, so that networks provide additional, higher-order
information on community structure. To date, studies of network diversity have
mostly been concerned with the distribution of interactions within locations,
and less so with variation among locations \citep{Dunne2006, Bascompte2007,
Ings2009, Kefi2012}. There is, however, ample evidence that interaction
networks vary in space and time \citep{Laliberte2010, Poisot2012, Albouy2014,
Poisot2016, Trojelsgaard2015}, even though there is no common framework with
which to generalize these results. Metacommunity theory provides explanations
for variation in the distribution of the nodes \citep{Gravel2011c,
Pillai2011}, but there is no such explanation to the variation of node and
link occurrences. Consequently, we urgently need a conceptual framework to
formalize these observations, as it is the only way towards fulfilling the
goal of community ecology: providing cogent predictions about, and
understanding of, the structure of ecological communities.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a clear need to bring the two together.
Here, we offer an integrated approach to do so, adopting the view that
community structure is best represented as a network of ecological
interactions. Based on this idea, we propose a new description of the basic
concept of the ecological niche that integrates the effect of the environment
on species distribution and on the ecological interactions among them.
Building on this concept, we develop a quantitative theory to explain turnover
of interactions in space and time. We first present the conceptual framework,
and then formalize it mathematically, using a probabilistic model to represent
the sampling of the regional pool of interactions. At the level of species
pairs, the statistical approach could be conceived as an interaction
distribution model. At the community level, the approach provides a
likelihood-based method to compare different hypotheses of network turnover.
As an illustrative example, we apply this novel framework to a large data set
on host–parasite interactions across Europe and find that two aspects of the
environment (temperature and precipitation) exert a strong imprint on species
co-occurrence, but not on species interactions. The network structure changes
systematically across the latitudinal gradient, with a peak of connectance at
intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, the \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized locally, revealing the stochastic nature of
ecological interactions. Our objective here is to provide a conceptual
framework to explain the sampling of the regional pool of interactions, along
with a quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-
occur locally and second why these species do or do not interact.

There are multiple causes of spatial turnover of species co-occurrence. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004a}. Stochasticity plays an additional role,
either because colonization and extinction events \citep{Hanski1999} are
inherently unpredictable or because strong non-linear feedbacks in community
dynamics generate alternative transients and equilibria \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. More recent approaches decompose total variation of the community
data into species and site contributions to beta diversity
\citep{Legendre2013}, and further partition it into dissimilarity due to
changes in species richness and dissimilarity due to actual species turnover
\citep{Baselga2010}. Even though these methods compare whole lists of species
among sites or measurements, they remain fundamentally "species-based", since
they report variation within columns. None of them explicitly considers
variation of associations (i.e., of pairs or higher-order motifs –
\citealt{Stouffer2007}).

Similarly, we are now getting a better understanding of interaction turnover. As
mentioned above, in the network approach to community structure, species and
interactions are represented by nodes and links, respectively. Associations
can also be represented by matrices in which entries represent the occurrence
or intensity of interactions among species (rows and columns). Network
complexity is then computed as the number of interactions (in the case of
binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition can be independent of variation of ecological
interactions, suggesting that species and interaction distribution may well
respond to different drivers \citep{Poisot2012}.

The "niche" is by far the dominant concept invoked to explain species
distributions and community assembly, from the local to the global scale.
Following \citealt{Hutchinson1957}, the niche is viewed as the set of
environmental conditions allowing a population to establish and persist (see
also \citealt{Holt2009a}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to put into practice with species distribution models, as it
maps naturally on available distributional and environmental data.
Consequently, a vast array of statistical tools have been developed to
implement it (e.g. BIOMOD \citealt{Thuiller2003}, MaxEnt
\citealt{Phillips2006}). It is however much harder to account for ecological
interactions within this approach \citep{Peterson2011}. As such, these
interactions are often viewed as externalities constraining or expanding the
range of environmental conditions required for a species to maintain a viable
population \citep{Pulliam2000, Soberon2007}.

Interestingly, the ecological network literature also has its own "niche
model" to position a species in a community \citep{Williams2000a}. The niche of
a species in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000a, Eklof2013}.
The niche model of food web structure and its variants have successfully
explained the complexity of a variety of networks, from food webs to
plant–pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}. This
conceptual framework is, however, limited to local communities, and does not
provide any explanation for the turnover of network structure along
environmental gradients.

@@

Despite several attempts to update the concept of the ecological niche,
ecologists have not moved far beyond the "n-dimensional hypervolume" defined
by Hutchinson. Despite its intuitive interpretation and easy translation into
species distribution models \citep{Boulangeat2012, Blonder2014}, the concept
has been frequently criticized \citep{Hardin1960, Peters1991, Silvertown2004},
and several attempts have been made to expand and improve it
\citep{Pulliam2000, Chase2003, Soberon2007, Holt2009a, McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}.
The Grinnellian dimension of the niche is the set of environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is the
effect of a species on its environment. While this aspect of the niche is well
known by community ecologists, it is trickier to turn into predictive models.
Nonetheless, the development of the niche model of food web structure
\citep{Williams2000a} and its parameterization using functional traits
\citep{Gravel2013, Bartomeus2016} made it more operational.

These perspectives are rather orthogonal to each other, and this has resulted
in considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
\emph{"[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions"}.
Their representation merges zero-net-growth isoclines delimiting the
Grinnellian niche ("when does the population persists?") with impact vectors
delimiting the Eltonian niche ("what is the per-capita impact?"). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}), it remains
impractical at larger spatial scales because of the difficulties in measuring
it. The absence of any mathematical representation of the niche that can be
easily fit to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions. Thus, a more integrative description of the niche
will be key to understand spatial and temporal turnover in community
structure.

@@

This approach radically changes the representation of the niche, putting
species distributions and ecological interactions into the same formalism.
Moreover, it allows the limits of the niche axes to be independent of each
other (as in the example in Fig. 2), or to interact. For instance, the optimal
prey size for predatory fishes could decline with increasing temperature
\citep{Gibert2014}, which would make diet boundaries functions of the
environment. Alternatively, we could also consider that the growth rate of the
predator changes with the size of its prey items, thereby altering the
environmental boundaries.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses concerning temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation, there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
constrain distribution \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent, and only the marginal occurrence must be accounted for, $P(
X_{ijy}|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1975,
Connor1979, Gotelli2000}). Second- and higher-order interactions (e.g. trophic
cascades) could also affect co-occurrence. Co-occurrence of multiple species
embedded in ecological networks is a topic of its own, however, and is
influenced by both network topology and species richness \citep{Cazelles2016}.
Not only direct interactions influence co-occurrence, but indirect
interactions do as well (e.g. plant species sharing an herbivore, or
herbivores sharing parasitoids, could repel each other in space
\citealt{Holt1993}). The impact of direct interactions and first-order
indirect interactions on co-occurrence tends to vanish with increasing species
richness in the community. Further, co-occurrence is also influenced by the
covariance of interacting species to an environmental gradient
\citep{Cazelles2015}. Because of the complexity of relating co-occurrence to
the structure of interaction networks, we will focus here on the variation of
interactions and not on their distribution, and leave this specific issue for
the Perspectives section and future research.

@@

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy} |E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to be sensitive to the environment. For instance,
\citep{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions also propagate up the
community and influence network structure \citep{Tylianakis2007, Woodward2010,
Petchey2010a}.

@@

Communities of willows, gallers, and parasitoids are species-rich and widely
distributed, with pronounced variation in community composition across space.
The genus Salix includes over 400 species, most of which are shrubs or small
trees \citep{Argus1997}, and is common in most habitats across the Northern
Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse community
of herbivorous insects, with one of the main herbivore groups being gall-
inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2003}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent "extended phenotypes" of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form, size,
and thickness of its gall.

@@

Our representation of spatial variation of community structure presents a new
approach for the study of the biogeography of ecological networks. We see the
following key challenges and opportunities ahead in this exciting area of
research:
